---
title: "ESM262_HW1_Kaapcke"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Tidy

1. Load tidyverse and read gazetteer data into a tibble. 

```{r tools_data, warning = F, message = F}

library(tidyverse)
library(knitr)

#Read in data
gaz_raw <- as_tibble(read_delim("C:/Users/lnkaa/Documents/ESM Computing/CA_Features_20180401.txt", delim = "|", col_names = T, col_types = cols(.default = col_character())))
#summary(gaz_raw)

```

2. Select desired columns.

```{r table, warning = F, message = F}

gaz <- gaz_raw %>% 
  select(FEATURE_ID:STATE_ALPHA, COUNTY_NAME, ends_with("_DEC"), ELEV_IN_M, MAP_NAME:DATE_EDITED)
#View(gaz)

```

3. Convert the columns to the appropriate type. Convert unknown placeholders to NA. 

```{r convert, warning = F, message = F}

#Replace "Unknown" with "NA"
is.na(gaz) <- gaz == "Unknown"

#Convert columns
gaz <- type_convert(gaz, col_types = cols( 
  FEATURE_ID = col_double(),
  PRIM_LAT_DEC = col_double(),
  PRIM_LONG_DEC = col_double(),
  SOURCE_LAT_DEC = col_double(),
  SOURCE_LONG_DEC = col_double(),
  ELEV_IN_M = col_double(),
  DATE_CREATED = col_date(format = "%m/%d/%Y"),
  DATE_EDITED = col_date(format = "%m/%d/%Y")
))
#Is there a way to convert the other columns features without knowing the number of levels? col_factor()??
#head(gaz) #Check conversion worked

```

4. Delete rows with NA for primary longitude or latitude or where the feature is not California.

```{r filter, message = F, warning = F}

gaz <- gaz %>%
  filter(STATE_ALPHA == "CA") %>% 
  filter(PRIM_LAT_DEC != 0, PRIM_LONG_DEC != 0)

```


5. Write gaz to a .csv

```{r write, message = F, warning = F}

# Get working directory and write the file
#getwd()
#write_delim(gaz, "gaz.csv", delim = "|")

#Why doesn't write_csv(gaz, "gaz.csv", delim = "|") work?

```

## Analyze the Data

1. What is the most frequenty appearing feature name?

```{r feature_name, message = F, warning = F}

gaz$FEATURE_NAME <- as.factor(gaz$FEATURE_NAME) # Type convert to a factor

# Count and arrange names in descending order
count_names <- gaz %>%
  group_by(FEATURE_NAME) %>% 
  summarize(n_names = n()) %>% 
  arrange(desc(n_names))

count_names

```

The feature name that appears the most often is "Church of Christ".

2. What is the least frequently occuring feature class?

```{r feature_class, message = F, warning = F}

# Type convert column to a factor
gaz$FEATURE_CLASS <- as.factor(gaz$FEATURE_CLASS)

# Count classes and arrange in ascending order
count_class <- gaz %>%
  group_by(FEATURE_CLASS) %>% 
  summarize(n_classes = n()) %>% 
  arrange(n_classes)

count_class

```

The feature classes that appear the least are "Sea" and "Isthmus".

3. What is the approximate center point of each county?

```{r county_center, warning = F, message = F}

# Find northernmost point
north <- gaz %>%
  group_by(COUNTY_NAME) %>% 
  summarize(northlat = max(PRIM_LAT_DEC))

# Find southernmost point
south <- gaz %>%
  group_by(COUNTY_NAME) %>% 
  summarize(southlat = min(PRIM_LAT_DEC))

# Join into a table
lats <- left_join(north, south, by = "COUNTY_NAME")

# Find westernmost point
west <- gaz %>%
  group_by(COUNTY_NAME) %>% 
  summarize(westlong = min(PRIM_LONG_DEC))

# Find easternmost point
east <- gaz %>%
  group_by(COUNTY_NAME) %>% 
  summarize(eastlong = max(PRIM_LONG_DEC))

# Join into a table
longs <- left_join(west, east, by = "COUNTY_NAME")

# Join into a table with all bounds
bounds <- left_join(lats, longs, by = "COUNTY_NAME")

# Create a new tibble with columns for the center latitude and longitude points.
center <- mutate(bounds,
                 Center_Latitude = southlat + ((northlat-southlat) / 2),
                 Center_Longitude = eastlong + ((westlong-eastlong) / 2)) %>% 
  select(COUNTY_NAME, Center_Latitude, Center_Longitude)
center

```

4. What are the fractions of the total number of features in each county that are natural vs. man-made?

```{r feature_type, message = F, warning = F}

classes <- gaz %>% 
  distinct(FEATURE_CLASS) %>% 
  arrange(FEATURE_CLASS)
View(classes)

types <- mutate(classes,
                Class_Type = c("Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Natural")
)
View(types)
  
  



```

center <- mutate(bounds,
                 Center_Latitude = southlat + ((northlat-southlat) / 2),
                 Center_Longitude = eastlong + ((westlong-eastlong) / 2)) %>% 
  select(COUNTY_NAME, Center_Latitude, Center_Longitude)


































